Stateless Model Checking with Data-Race Preemption Points

### 2015-10-02 pldi-16/intro.tex
split pldi multiple tex files

%
% LaTeX template for prepartion of submissions to PLDI'15
%
% Requires temporary version of sigplanconf style file provided on
% PLDI'15 web site.
%
\documentclass[pldi]{sigplanconf-pldi15}

%
% the following standard packages may be helpful, but are not required
%
\usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI

\begin{document}

%
% any author declaration will be ignored  when using 'plid' option (for double blind review)
%

% TODO: Don't forget to run e.g. s/\\quicksand /\\quicksand~/
\newcommand\landslide{\textsc{Landslide}}
\newcommand\quicksand{\textsc{Quicksand}}

\title{Stateless Model Checking with Data-Race Preemption Points}
\authorinfo{Ben Blum}{Carnegie Mellon University}{bblum@cs.cmu.edu}
\authorinfo{Garth Gibson}{Carnegie Mellon University}{garth@cs.cmu.edu}

\maketitle
\begin{abstract}
Stateless model checking is a promising technique for testing concurrent programs,
but is vulnerable to exponential explosion of the state space when the test input parameters are too large.
Several partial-order reduction techniques exist for mitigating this explosion,
but even after pruning equivalent interleavings, the state space size for a fixed set of preemption points is unpredictable and often intractable.
%
Data race detection, another concurrency testing approach, focuses on identifying suspicious memory access pairs during a single test execution.
It avoids concerns of intractable state space size, but suffers from a high rate of false positives.

We present an {\em Iterative Deepening} framework for stateless model checking,
which manages the exploration of many state spaces using different subsets of preemption points.
It uses state space estimation to prioritize jobs most likely to complete in a fixed CPU budget,
and it incorporates a data-race analysis to dynamically add new preemption points to verify each data race as buggy or benign.
%
Our evaluation shows this technique is
more effective than single-state-space model checking, both at finding more bugs and at completing more state spaces when no bug exists.

\end{abstract}

@@

% Can I even make a claim this broad to begin with?
However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
For example, \textsc{CHESS} \cite{chess} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
It provides an additional data-race analysis to report any violations of this model;
% TODO: Make sure that there are enough citations for this claim.
however, data-race analyses are prone to report false positives and benign races which require annotations or imprecise heuristics to reduce \cite{racerx,tsan,datacollider}.
%
On the other hand,
% TODO: Find a systematic tester that does this!
\textsc{Ski} \cite{ski}
is able to preempt threads around any shared resource access. Such fine granularity would automatically check if each data race is a real bug, but makes full state space completion intractable for all but the most rudimentary test inputs.
%
This work shows how to avoid making that tradeoff decision in advance.

We present \quicksand, a framework for {\em Iterative Deepening} of preemption points during stateless model checking.
Named after the analogous technique in chess AI \cite{iterative-deepening-chess}, our approach likewise makes progressively deeper searches of the state space until a given CPU budget is exhausted.
Rather than attempting to search a single state space with every preemption point among a fixed set enabled (e.g., preempting on every pthread API call),
\quicksand~tests many different state spaces corresponding to subsets of those points, managing a model checker instance to explore each one.
It estimates the size of each state space to decide when long-running instances should be suspended, and dynamically generates new state spaces based on data race analysis.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Design}

TODO

Unlike in chess AI, however, adding more preemption points to a completed state space is not necessarily a linear process, as  % blah blah multiple combinations

% define minimal state space and maximal state space

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Implementation}

Landslide provides estimated completion time using the Recursive Estimator algorithm in \cite{estimation}.

% TODO: Talk about the multiple ways DR PPs are added (alone / in addition)

% TODO: make sure to talk about avoiding thrashing
\subsection{Avoiding thrashing in \quicksand}

(Here the reviewers will wonder if the whole point of the paper was just to build up to that one joke.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Evaluation}

% TODO: Don't forget to add to quicksand

Evaluation questions:
\begin{enumerate}
	\item Easy question: Is it good to incorporate DR analysis for new PPs? (Do we find DR-PP-only bugs?)
	\item How does this compare to static DR detection (how many DRs does ID verify as benign?)
	\item DRs aside, is ID more effective than the maximal state space? (Do we find bugs in subset spaces, when the max state space would time out?)
		% Make sure to argue how this shows ID is applicable to e.g. message-passing SMC tools. "Even when DR PPs aren't on the table, you should do our thing."
	\item Is ID useful for managing new DR PPs? (Are there some DRs where the PP alone finds the bug, and others where the old PP plus the new DR PP is needed?)
		% How many DR bugs were found from DR-alone state space, and how many from sema_foo + DR state space? What we want is a *variety*, which proves that iterative deepening is better than either individual strategy.
	\item How effective is the free-re-malloc technique at eliminating false positives? (Just turn the optimization off, and measure how many more SSes are generated.)

% Figure out concretely what the data race tricks are that we do, so we can claim them as contributions in the paper. Then ACTUALLY EVALUATE THEM.
%         - Speculative DR PPs.
%                 Not a heuristic, rather how to make it work at all to begin with.
%                 (Cite MS thesis, claim on backwards explorating finding bugs faster)
%         - Free/re-malloc to eliminate some false positives. See #193.
%                 Measure how many false positives are eliminated.
%                 Check, ofc, to make ABSOLUTE SURE, that no bugs missed w/ this trick.
%                         If there are, it could be because of the implementation
%                         bug described in #193.
%         - Using tid/last_call filtering because whole stack traces are too expensive.
%                 Turn on/off and measure how resulting DR bug #s change.
%         - Optional: Reprioritizing DRs based on "confirmed" / "suspected"
%                 Shouldn't be hard just make ID wrapper print "s" or "c"!
%                 Is it helpful for ID to put priorities on DR PPs?
%                         Test by inverting the priority and see if fewer buges are found.
%         // Super optional to talk about. Probably not worth the time.
%         // - "Too suspicious" (during init/destroy)
%         //      (Cite eraser, section 2.2)

%% Notes from jiri meeting. Some overlap with stuff above.
% Evaluation in 2 major questions
%         Data races (Sequence this first) vs no-DR-systesting
%                 Sub questions: Compare to single-pass DRs
%                         What false positive rate are we avoiding by comparison?
%                         How often would the DR not even be caught on the 1st execution, requiring backtracks to expose?
%                                 (don't forget set EXPLORE_BACKWARDS=0 here)
%         "Control experiment vs ID" (PLDI cares less asbout "finding bugs faster")

\end{enumerate}

We tested 87 pintos kernels and found races in 45-47 of them, among which 34-47 of those bugs required data-race preemption points to expose. That means we rock.

% TODO: Future work.
% Future work: Add a way to configure even smaller subsets (eg "only mutex_locks called from site X") for cases where mx_lock and mx_unlock alone are still too big. Count the number of kernels for which this was the case.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}

%None. We are the first ever to explore this area, and of course expect the Turing Award for our groundbreaking insights.

\quicksand~is built upon many established concurrency testing techniques, dating back of course to Verisoft, the original C model checker \cite{verisoft}.
We use \landslide~as our model checker \cite{landslide},
% TODO: is DPOR abbreviated earlier in the paper
which itself implements DPOR \cite{dpor},
state space estimation \cite{estimation},
and data race detection \cite{eraser}. Related model checkers can largely be classified based on their treatment of shared-memory thread communication.

{\bf Synchronization events only.} CHESS \cite{chess} and \textsc{dBug} \cite{dbug-ssv} instrument the thread library API, and can preempt programs only during calls to this API.
Hence they will miss any crashes resulting from interleaving threads at instruction granularity during a data race. CHESS provides a data-race analysis to report any such violations of its concurrency model to the user, but does not incorporate data race candidates as PPs in future tests.

{\bf Message-passing.} Other stateless model checkers, such as SAMC \cite{samc}, \textsc{MaceMC} \cite{macemc}, \textsc{MoDist}, and ETA \cite{parallel-dpor}, require thread communication to be restricted to a message-passing API to more effectively test distributed systems.
This eliminates the need for data race analysis, but restricts the class of programs that can be tested.
Nevertheless, Iterative Deepening is still applicable to these tools.

{\bf Preempting at instruction granularity} is a prerequisite for using data-race PPs. However, every such prior tool we know of has serious drawbacks.
SKI \cite{ski} approaches kernel code by choosing in advance a random set of instruction offsets from the start of the test,
which is more similar to stress testing or fuzzing than to exhaustive state space exploration.
% TODO: Make sure you talk about your mutex experiment, or rephrase this.
SPIN \cite{spin} specializes in verifying synchronization primitive implementations such as RCU, which is very similar to our mutex-testing experiment.
However, it does not employ Iterative Deepening, and requires programs to be written in the PROMELA DSL, so cannot check implementations directly.

{\bf Other techniques.} Various improvements to DPOR have been proposed, such as Dynamic Interface Reduction \cite{demeter}, Maximal Causality Reduction \cite{mcr}, and DPOR for TSO/PSO \cite{tsopso}.
These are all orthogonal to our technique.
\textsc{Parrot} \cite{parrot} combines model checking with a partially-determinizing runtime environment, but still, fewer than half the non-trivial state spaces in their evaluation could be completed, providing a strong argument for \quicksand.
Finally, Iterative Context Bounding \cite{chess} is most similar to Iterative Deepening, as both approaches provide a concrete partial verification on some subset of interleavings when full completion is intractable.
However, ICB still limits itself to a fixed set of PPs, and so far no algorithm has been proposed to dynamically add data race PPs during a test with ICB.

% TODO: talk about data race detectors???

% TODO: Mention dbug+PARROT as a motivating argument for our work. The number of tests with INF state space, EVEN WITH ONLY PTHREAD API PPs, shows that fixed-in-advance-pp-set needs our work to improve.
%Indeed, the low state space completion rate in dBug+Parrot's evaluation \cite{parrot} shows that for some tests even the set of all pthread API calls is too many PPs.

% TODO: Address ICB. Note that BPOR paper claims that ICB(3+) repeats LOADS of work, and that makes it ok for landslide-ID to repeat work.

% IDK if i should mention it, but OOPSLA 2015, protocol based verification of MPI concurrency paper. Different verification approach entirely; doesn't suffer exponential explosion but limited to programs with no shared state and MPI communication only

% Probably NOT worth a mention: OOPSLA 2015, stateless model checking of event driven applications. Turning timer-driven model on its head and checking single-threaded, but asynch-event-driven programs (i.e. device-like signal handlers)

% TODO: Read OOPSLA 2015 "SATcheck, sat-directed SMC for SC/TSO"

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

We are
%overwhelmingly
great. Accept our paper.

% TODO: Publish source of ID wrapper alone, if can't publish landslide itself.
% TODO: Publish log files.

\section{Acknowledgements}

Thanks.

\bibliographystyle{abbrvnat}
\bibliography{citations}{}

\end{document}

>>>>

@@

% Can I even make a claim this broad to begin with?
However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
% TODO: You repeat the following 1.5 sentence in the related work section; maybe you can eliminate this redundancy to save space?
For example, \textsc{CHESS} \cite{chess} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
It provides an additional data-race analysis to report any violations of this model;
however, data-race analyses are prone to report false positives and benign races which require annotations or imprecise heuristics to reduce \cite{racerx,tsan,datacollider}.
%
On the other hand, SPIN \cite{spin}
is able to preempt threads around any shared memory access. Such fine granularity would automatically check if each data race is a real bug, but makes full state space completion intractable for all but the most rudimentary test inputs.
%
This work shows how to avoid making that tradeoff decision in advance.

We present \quicksand, a framework for {\em Iterative Deepening} of preemption points during stateless model checking.
Named after the analogous technique in chess AI \cite{iterative-deepening-chess}, our approach likewise makes progressively deeper searches of the state space until a given CPU budget is exhausted.
Rather than attempting to search a single state space with every available preemption point enabled (e.g., preempting on every pthread API call),
\quicksand~tests many different state spaces corresponding to subsets of those points, managing a model checker instance to explore each one.
It estimates the size of each state space to decide when long-running instances should be suspended, and dynamically generates new state spaces based on data race analysis.

@@

### 2015-10-02 pldi-16/paper.tex
split pldi multiple tex files

\title{Systematic Testing with Data-Race Preemption Points}
\authorinfo{Ben Blum}{Carnegie Mellon University}{bblum@cs.cmu.edu}
\authorinfo{Garth Gibson}{Carnegie Mellon University}{garth@cs.cmu.edu}

@@

\section{Introduction}

% blah blah trite opening sentence
%As parallelism becomes ever more important for achieving high performance in modern-day programs,
%so too do advanced concurrency testing techniques become important for verifying the correctness of those programs.
Concurrency bugs are notoriously hard to find and reproduce because they only appear in specific thread interleavings, which arise at random during normal program execution.
{\em Stateless model checking} \cite{verisoft} offers a method for finding such bugs,
or verifying their absence,
%by systematically executing a program along as many distinct interleavings as possible,
by forcing a program to execute each distinct interleaving,
capturing and controlling this nondeterminism in a finite state space.
Unfortunately, these state spaces explode exponentially in the size of the input program.
Reduction techniques such as Dynamic Partial Order Reduction (DPOR) \cite{dpor} and Maximal Causality Reduction (MCR) \cite{mcr} expand the limits of feasible test completion,
and search ordering strategies such as Iterative Context Bounding (ICB) \cite{chess} allow bugs to be found sooner in a given exploration should the exist.

% Can I even make a claim this broad to begin with?
However, all stateless model checkers to date are bound by a fixed set of {\em preemption points} (PPs): code locations that define the granularity at which threads interleave.
For example, CHESS \cite{chess} preemptions only on synchronization operations and library calls, which can miss lock-free shared memory races.
It provides an additional data-race analysis to report any violations of this model;
% TODO: Make sure that there are enough citations for this claim.
however, data-race analyses are prone to report false positives and benign races which require annotations or imprecise heuristics to reduce \cite{racerx,tsan,datacollider}.
%
On the other hand,
% TODO: Find a systematic tester that does this!
SKI \cite{ski}
is able to preempt threads around any shared resource access. Such fine granularity would automatically check if each data race is a real bug, but makes state space completion intractable for all but the most rudimentary test inputs.
%
This work shows how to avoid making this tradeoff decision in advance.

\newcommand\landslideid{\textsc{Landslide-ID}}
We present \landslideid...

\section{Implementation}

Landslide provides estimated completion time using the Recursive Estimator algorithm in \cite{estimation}.

% TODO
TODO

% TODO: make sure to talk about avoiding thrashing

\section{Evaluation}

Evaluation questions:
\begin{enumerate}
	\item Is ID more effective than 
\end{enumerate}

We tested 87 pintos kernels and found races in 45-47 of them, among which 34-47 of those bugs required data-race preemption points to expose. That means we rock.

% TODO: Future work.
% Future work: Add a way to configure even smaller subsets (eg "only mutex_locks called from site X") for cases where mx_lock and mx_unlock alone are still too big. Count the number of kernels for which this was the case.

\section{Related Work}

None. We are the first ever to explore this area, and of course expect the Turing Award for our groundbreaking insights.

% dbug-like systematic testers: dbug, chess, modist

% SKI-like systematic testers: SKI

% other: macemc(???), ETA (message passing only)

% TODO: Mention dbug+PARROT as a motivating argument for our work. The number of tests with INF state space, EVEN WITH ONLY PTHREAD API PPs, shows that fixed-in-advance-pp-set needs our work to improve.

@@

Thanks.

>>>>

% TODO: Don't forget to run e.g. s/\\quicksand /\\quicksand~/
\newcommand\landslide{\textsc{Landslide}}
\newcommand\quicksand{\textsc{Quicksand}}

\title{Stateless Model Checking with Data-Race Preemption Points}
\authorinfo{Ben Blum}{Carnegie Mellon University}{bblum@cs.cmu.edu}
\authorinfo{Garth Gibson}{Carnegie Mellon University}{garth@cs.cmu.edu}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{intro}
\input{di}
\input{evaluation}
\input{related}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Many thanks to David A. Eckhardt, Vaishaal Shankar, and Haryadi Gunawi for generously providing student implementations from CMU's, Berkeley's, and U. of Chicago's OS classes respectively.
Thanks to Ji\v{r}\'{i} \v{S}im\v{s}a and the anonymous PLDI reviewers for their helpful comments.
This work was supported in part by % TODO: funding sources
